---
title: "Wrangle pangolin data"
output: html_document
---

This script takes georeferenced data and IUCN range maps for pangolins and wrangles these for further analysis. sf objects are hard to export, so we suggest after running this script you move straight to plotting and analysis scripts. Or run these via sourcing these at the bottom of this script.

## Load libraries
```{r, message = FALSE, warning = FALSE}
## Install sfe from GitHub
##devtools::install_github("JCur96/sfe", force = TRUE)

library(sfe)
library(rgeos)
library(rgdal)
library(Hmisc)
library(tidyverse)
library(here)
library(patchwork)

# Short function to get decades
floor_decade <- function(x){
  if(!is.na(x)){
  x - x %% 10
  }else{NA_character_}
}
```

## Read in the georeferenced data
```{r, message = FALSE, warning = FALSE}
## NHM georeferenced data
nhm <- read_csv(here("raw-data/NHM_pangolins_June2019.csv"))

# Remove material column and unknown species names or port specimens
nhm <- 
  nhm %>% 
  select(-Material) %>%
  filter(Binomial != "unknown") %>%
  # Remove specimens from ports
  filter(RegistrationNumber != "12.12.3.3" & RegistrationNumber != "no reg 9") %>%
  # Remove specimens from Sudan as this must be an error.
  filter(RegistrationNumber != "87.12.1.103" & RegistrationNumber != "7.12.1.102")

## GBIF georeferenced data
gbif <- read_csv(here("raw-data/gbif_pangolins_Oct2019.csv"))

## Exclude the specimens which are either misidentified or have incorrect localities
## These are Phataginus species founr in southern Africa where the genus is not found
## And very far from their current range. Could be Smutsia, but we can't check the specimens
gbif <-
  gbif %>%
  filter(RegistrationNumber != "BOUM_11.Man.1" & 
         RegistrationNumber != "MVZ_MVZ:Mamm:4829" &
         RegistrationNumber != "AMNH_M-216259" &
         RegistrationNumber != "SUI_18440" &
         RegistrationNumber != "O_18156" &
         RegistrationNumber != "GNM_Ma ex 1292" &
         RegistrationNumber != "GNM_CollAn 7491")
```

## Read in the IUCN range maps
```{r, message = FALSE, warning = FALSE}
iucn <- readOGR(dsn = here("raw-data/IUCN_pholidota"), 
                layer = 'maps_pholidota')
```

## Read in the ecology data
```{r, message = FALSE, warning = FALSE}
ecology <- read_csv(here("raw-data/pangolin-ecology.csv"))
```

## Add/Merge all the georeferenced and ecology data together
```{r}
specs <- 
  nhm %>%
  rbind(gbif) %>%
  full_join(ecology, by = c("Binomial" = "binomial"))
```

## Identify duplicates

- these are either specimens from the same species with the exact same latitude and longitude coordinates or...
- specimens from the same species with the exact same latitude and longitude coordinates and the same year
- adds two columns identifying duplicates to exclude in analyses

```{r}
specsx <-
  specs %>%
  dplyr::select(Binomial, Latitude, Longitude)

specsx_year <-
  specs %>%
  dplyr::select(Binomial, Latitude, Longitude, Year)

specs <- 
  specs %>%
  mutate(duplicates = duplicated(specsx)) %>%
  mutate(duplicates_year = duplicated(specsx_year))
```
## Tidy the georeferenced data

- Change species binomial columns to "binomial" as many of the functions rely on the data having a column of species names under the column header of "binomial". 
- Filter out blank or NA values from binomial, latitude and longitude. 
- Convert to the sf format, ensuring correct projection in latitude and longitude.
  
```{r}
## Tidy georeferenced data
specs <- 
  specs %>%
  rename(binomial = Binomial) %>%
  # Exclude specimens with no name or coordinates
  filter(!is.na(binomial) & binomial != "unknown" &
         !is.na(Latitude) & !is.na(Longitude)) %>%
  # Add decade variable (function above)
  # This maps to character to deal with NAs so needs coercing back to numeric
  mutate(Decade = map_chr(Year, floor_decade)) %>%
  mutate(Decade = as.numeric(Decade)) %>%
  # Convert extents to km
  mutate(Extent_km = Extent_m / 1000) %>%
  # Remove non needed columns
  dplyr::select(-NOTES, -LocalityType, -Locality, -Extent_m) %>%
  # Convert to sf format
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)
```

## Tidy the IUCN range maps

- Convert to the sf format, ensuring correct projection in latitude and longitude.
```{r}
iucn <- 
  iucn %>%
  st_as_sf() %>%
  st_transform(4326)
```

## Add error polygons to data

- change the Extent column into kilometers and rename it
- add a `buffer` to the point data using the point-radius method.
```{r, message = FALSE, warning = FALSE}
specs_errors <-
  specs %>%
  addError()
```

## Calculate "overlaps" and "centroid-distances"
The next set of functions calculate percentage overlap between point-radius data and IUCN range maps for each species. Then extract the binomial overlap or not. Then add distance column (and a distance2 column, as there are a few IUCN records with two range polygons, and
without this inclusion the function will not run).

```{r, message = FALSE, warning = FALSE}
# For speed and viewability we strip out all the extraneous stuff,
# we just need the species name and the geometry
myvars <- c('binomial', 'geometry')
iucn2 <- iucn[myvars]
iucn2 <- resolveIUCNGeom(iucn2)
iucn2 <- st_as_sf(iucn2)

# Work out the percentage overlaps among point-radius polygons and IUCN ranges
overlaps <- calculateOverlaps(specs_errors, iucn2)
# Flatten this to 0 for 0% overlap, and 1 for > 0% overlap
overlaps <- binomialOverlap(overlaps)
```

We can now use `overlaps` in our analyses and plots. 

## Plot maps
```{r}
source(here("analyses/02-figure-maps.R"))
```


