---
title: "Create .shp files and calculate overlaps"
output: html_document
---

Load necessary libraries
```{r}
library(rgdal)
library(sf)
library(stars)
library(sfe)
```

Convert .tif files made using GIS software (ARC) into .shp files
.tifs/.shp files are stored on the NHM data portal as they are too large to 
practically upload and host on git.

```{r}
GeoTiffToPolygon <- function(directory) 
{
  ##get the full path as a string
  fileList = list.files(path = directory, pattern = "*.tif", full.names = TRUE)
  for (file in fileList) 
  {
    specificName = str_extract(file, regex("\\/{1}\\w+\\."))
    specificName = gsub("/", "", specificName)
    specificName = gsub("\\.", "\\.shp", specificName)
    ##pass to read_stars
    rasterFile = read_stars(file)
    ##pass to st_as_sf
    polygonFile = st_as_sf(rasterFile)
    ##check to see if outut dir exists and create if not
    dir.create(file.path("../data/shpFiles"), showWarnings = FALSE)
    ##save the new object
    st_write(polygonFile, paste0("../data/shpFiles", "/", specificName))
  }
}

# not sure how Natalie has dir set up so e.g.
GeoTiffToPolygon("../data/aohRaw/")
```

Overlap calculation functions needed rewriting as AOH .shp files are very large
and old functions were not memory efficient enough to deal with them.
```{r}
#Functions go here
# two input function for calculating the percentage overlap
calcOverlaps <- function(df1, df2) {
  df1 <- lwgeom::lwgeom_make_valid(df1)
  df2 <- lwgeom::lwgeom_make_valid(df2)
  # gives percentage overlap between NHM and IUCN
  overlap <- st_intersection(df1, df2) %>% st_area() / st_area(df1, df2) * 100
  # at this point the output is of class "units" which don't play nice
  overlap <- units::drop_units(overlap)
  # allows for handling of cases of zero overlap
  if (purrr::is_empty(overlap) == T) {
    # as it otherwise returns a list of length zero,
    # which cannot be appended to a df
    overlap <- c(0)
  }
  overlap <- as.list(overlap)
  # returns the result, so can be passed to another fun
  return(overlap)
}

hullOverFun <- function(df1, df2) {
  df1$Percent_overlap <- NA
  for (row in 1:nrow(df1)) {
    # extract the geometry
    geom <- df1$geometry[row]
    x <- calcOverlaps(geom, df2)
    df1$Percent_overlap[row] <- x
  }
  return(df1) # return the modified df for use in another fun
}

calculateOverlaps <- function(NHM, AOH) {
  # create an empty list to store results
  output <- c()
  NHM <- st_transform(NHM, 2163)
  AOH <- st_transform(AOH, 2163)
  tmp <- hullOverFun(NHM, AOH)
  # rebuilding the input df with a new col
  output <- rbind(tmp, output)
  output$Percent_overlap <- as.numeric(output$Percent_overlap)
  return(output)
}
MakeNHMIntoDFList <- function(NHM_Pangolins) 
{
  pangolinDfList <- split(NHM_Pangolins, f = NHM_Pangolins$binomial)
  return(invisible(pangolinDfList))
}
RunAOHAnalysis <- function(NHM_Pangolins, AOHDataDir, isIUCN = F)
{
  NHMPangolinList <- MakeNHMIntoDFList(NHM_Pangolins)
  AOHFileList = list.files(path = AOHDataDir, pattern = "*.shp", full.names = TRUE)
  for (file in AOHFileList)
  {
    sppName = str_extract(file, regex("\\/\\w+\\_\\w+"))
    sppName = gsub("/", "", sppName)
    sppName = gsub("\\_$", "", sppName)
    # read the file in!
    sppFile = st_read(dsn = file)
    if (isIUCN == T) {
      # change BINOMIAL to binomial
      names(file)[names(file) == "BINOMIAL"] <- "binomial"
    }
    else
    {
      names(sppFile)[1] <- 'binomial'
      sppName <- gsub("_aoh", "", sppName)
    }
    print(sppName)
    # pass it to this little pipline
    sppFile$binomial <- sppName
    sppFile <-
      sppFile %>%
      st_as_sf() %>%
      st_union()
    sppFile <- st_make_valid(sppFile)
    # calculate the overlaps and append to a df
    for (item in NHMPangolinList) {
      Spp <- unique(item$binomial)
      if (unique(item$binomial) == sppName) {
        overlaps <- calculateOverlaps(item, sppFile)
        overlaps <- binomialOverlap(overlaps)
        fullFilePath <- paste("../data/overlaps_", sppName, ".csv", sep = "")
        if (file.exists(fullFilePath)) { #if you want new output/ fresh rerun must delete all output files
          print('file already exists, this run is likely for just land use!')
          newFileName <- paste("../data/overlaps_", sppName, "_landUse", ".csv", sep = "")
          st_write(overlaps, newFileName)
        }
        if (!file.exists(fullFilePath)) {
          st_write(overlaps, paste("../data/overlaps_", sppName, ".csv", sep = ""))
        }
      }
    }
  }
}
```

Run the .shp files through these new functions, with the output generated in the
previous script (specs object)
```{r}
RunAOHAnalysis(specs, "../data/shpFiles")
```
Output stored in X???????