---
title: "HPD change analyses"
output: html_document
---

## Load libraries
```{r, message = FALSE, warning = FALSE}
library(sf)
library(sfe)
library(tidyverse)
library(here)
library(broom)
library(ggfortify)
library(betareg)
library(lmtest)
```

## Prepare for the analyses
First read in the data as CSV not spatial data as we don't need the spatial geometries here. Relevel redlist so it matches increasing extinction risk.
```{r, message = FALSE, warning = FALSE}
overlaps_hpd <- 
  read_csv(here("data/overlaps_hpd.csv")) %>%  
  mutate(redlist = factor(redlist, levels = c("VU", "EN", "CR")))
```

Next read in overlaps data so we can get the number of overlaps
```{r}
overlaps <- 
  read_csv(here("data/overlaps_combined.csv")) 

overlaps <- 
  as.data.frame(overlaps) %>%
  dplyr::select(RegistrationNumber, Percent_overlap)

overlaps_hpd <- left_join(overlaps_hpd, overlaps, 
                              by = "RegistrationNumber")
```

## Models - with % area overlaps
Beta regression with logit link

First rescale percent to 0-1
```{r}
overlaps_hpd$Percent_overlap <- overlaps_hpd$Percent_overlap / 100
```

Next rescale to remove zeros and 1s
```{r}
transform01 <- function(x) {
  (x * (length(x) - 1) + 0.5) / (length(x))
}

overlaps_hpd$Percent_overlap_scaled <- 
  transform01(overlaps_hpd$Percent_overlap)
```

Variables are:
* popc = population count
* popd = population density

Note that the population changes are huge and skewed so will need to be log transformed, but a small number are negative, so first we need to sort that out...

```{r}
fix.min <- function(x){
  x + abs(min(x, na.rm = TRUE)) + 1
}

overlaps_hpd <-
  overlaps_hpd %>%
  mutate(log_popc_1850 = log(fix.min(popc_1850))) %>%
  mutate(log_popc_1900 = log(fix.min(popc_1900))) %>%
  mutate(log_popc_1950 = log(fix.min(popc_1950))) %>%
  mutate(log_popd_1850 = log(fix.min(popd_1850))) %>%
  mutate(log_popd_1900 = log(fix.min(popd_1900))) %>%
  mutate(log_popd_1950 = log(fix.min(popd_1950)))
```

## Population Count
### popc_1850
```{r}
# Fit model with fixed phi
model1a1 <- betareg(Percent_overlap_scaled ~ log_popc_1850, data = overlaps_hpd)
# Fit model with variable phi
model1a2 <- betareg(Percent_overlap_scaled ~ log_popc_1850 | log_popc_1850, data = overlaps_hpd)

# Which is better fit?
lrtest(model1a1, model1a2)

```

Check results are the same with bias correction and bia reduction
```{r}
# Fit model with fixed phi
model1a1.bc <- betareg(Percent_overlap_scaled ~ log_popc_1850, data = overlaps_hpd, type = "BC")
# Fit model with variable phi
model1a2.bc <- betareg(Percent_overlap_scaled ~ log_popc_1850 | log_popc_1850, data = overlaps_hpd, type = "BC")
# Which is better fit?
lrtest(model1a1.bc, model1a2.bc)
```

```{r}
# Fit model with fixed phi
model1a1.br <- betareg(Percent_overlap_scaled ~ log_popc_1850, data = overlaps_hpd, type = "BR")
# Fit model with variable phi
model1a2.br <- betareg(Percent_overlap_scaled ~ log_popc_1850 | log_popc_1850, data = overlaps_hpd, type = "BR")

# Which is better fit?
lrtest(model1a1.br, model1a2.br)
```

```{r}
## Model outputs
summary(model1a2)
```

### popc_1900
```{r}
# Fit model with fixed phi
model1a1 <- betareg(Percent_overlap_scaled ~ log_popc_1900, data = overlaps_hpd)
# Fit model with variable phi
model1a2 <- betareg(Percent_overlap_scaled ~ log_popc_1900 | log_popc_1900, data = overlaps_hpd)

# Which is better fit?
lrtest(model1a1, model1a2)

```

Check results are the same with bias correction and bia reduction
```{r}
# Fit model with fixed phi
model1a1.bc <- betareg(Percent_overlap_scaled ~ log_popc_1900, data = overlaps_hpd, type = "BC")
# Fit model with variable phi
model1a2.bc <- betareg(Percent_overlap_scaled ~ log_popc_1900 | log_popc_1900, data = overlaps_hpd, type = "BC")
# Which is better fit?
lrtest(model1a1.bc, model1a2.bc)
```

```{r}
# Fit model with fixed phi
model1a1.br <- betareg(Percent_overlap_scaled ~ log_popc_1900, data = overlaps_hpd, type = "BR")
# Fit model with variable phi
model1a2.br <- betareg(Percent_overlap_scaled ~ log_popc_1900 | log_popc_1900, data = overlaps_hpd, type = "BR")

# Which is better fit?
lrtest(model1a1.br, model1a2.br)
```

```{r}
## Model outputs
summary(model1a2)
```

### log_popc_1950
```{r}
# Fit model with fixed phi
model1a1 <- betareg(Percent_overlap_scaled ~ log_popc_1950, data = overlaps_hpd)
# Fit model with variable phi
model1a2 <- betareg(Percent_overlap_scaled ~ log_popc_1950 | log_popc_1950, data = overlaps_hpd)

# Which is better fit?
lrtest(model1a1, model1a2)

```

Check results are the same with bias correction and bia reduction
```{r}
# Fit model with fixed phi
model1a1.bc <- betareg(Percent_overlap_scaled ~ log_popc_1950, data = overlaps_hpd, type = "BC")
# Fit model with variable phi
model1a2.bc <- betareg(Percent_overlap_scaled ~ log_popc_1950 | log_popc_1950, data = overlaps_hpd, type = "BC")
# Which is better fit?
lrtest(model1a1.bc, model1a2.bc)
```

```{r}
# Fit model with fixed phi
model1a1.br <- betareg(Percent_overlap_scaled ~ log_popc_1950, data = overlaps_hpd, type = "BR")
# Fit model with variable phi
model1a2.br <- betareg(Percent_overlap_scaled ~ log_popc_1950 | log_popc_1950, data = overlaps_hpd, type = "BR")

# Which is better fit?
lrtest(model1a1.br, model1a2.br)
```

```{r}
## Model outputs
summary(model1a2)
```

## Population density
### log_popd_1850
```{r}
# Fit model with fixed phi
model1a1 <- betareg(Percent_overlap_scaled ~ log_popd_1850, data = overlaps_hpd)
# Fit model with variable phi
model1a2 <- betareg(Percent_overlap_scaled ~ log_popd_1850 | log_popd_1850, data = overlaps_hpd)

# Which is better fit?
lrtest(model1a1, model1a2)

```

Check results are the same with bias correction and bia reduction
```{r}
# Fit model with fixed phi
model1a1.bc <- betareg(Percent_overlap_scaled ~ log_popd_1850, data = overlaps_hpd, type = "BC")
# Fit model with variable phi
model1a2.bc <- betareg(Percent_overlap_scaled ~ log_popd_1850 | log_popd_1850, data = overlaps_hpd, type = "BC")
# Which is better fit?
lrtest(model1a1.bc, model1a2.bc)
```

```{r}
# Fit model with fixed phi
model1a1.br <- betareg(Percent_overlap_scaled ~ log_popd_1850, data = overlaps_hpd, type = "BR")
# Fit model with variable phi
model1a2.br <- betareg(Percent_overlap_scaled ~ log_popd_1850 | log_popd_1850, data = overlaps_hpd, type = "BR")

# Which is better fit?
lrtest(model1a1.br, model1a2.br)
```

```{r}
## Model outputs
summary(model1a2)
```

### log_popd_1900
```{r}
# Fit model with fixed phi
model1a1 <- betareg(Percent_overlap_scaled ~ log_popd_1900, data = overlaps_hpd)
# Fit model with variable phi
model1a2 <- betareg(Percent_overlap_scaled ~ log_popd_1900 | log_popd_1900, data = overlaps_hpd)

# Which is better fit?
lrtest(model1a1, model1a2)

```

Check results are the same with bias correction and bia reduction
```{r}
# Fit model with fixed phi
model1a1.bc <- betareg(Percent_overlap_scaled ~ log_popd_1900, data = overlaps_hpd, type = "BC")
# Fit model with variable phi
model1a2.bc <- betareg(Percent_overlap_scaled ~ log_popd_1900 | log_popd_1900, data = overlaps_hpd, type = "BC")
# Which is better fit?
lrtest(model1a1.bc, model1a2.bc)
```

```{r}
# Fit model with fixed phi
model1a1.br <- betareg(Percent_overlap_scaled ~ log_popd_1900, data = overlaps_hpd, type = "BR")
# Fit model with variable phi
model1a2.br <- betareg(Percent_overlap_scaled ~ log_popd_1900 | log_popd_1900, data = overlaps_hpd, type = "BR")

# Which is better fit?
lrtest(model1a1.br, model1a2.br)
```

```{r}
## Model outputs
summary(model1a2)
```

### log_popd_1950
```{r}
# Fit model with fixed phi
model1a1 <- betareg(Percent_overlap_scaled ~ log_popd_1950, data = overlaps_hpd)
# Fit model with variable phi
model1a2 <- betareg(Percent_overlap_scaled ~ log_popd_1950 | log_popd_1950, data = overlaps_hpd)

# Which is better fit?
lrtest(model1a1, model1a2)

```

Check results are the same with bias correction and bia reduction
```{r}
# Fit model with fixed phi
model1a1.bc <- betareg(Percent_overlap_scaled ~ log_popd_1950, data = overlaps_hpd, type = "BC")
# Fit model with variable phi
model1a2.bc <- betareg(Percent_overlap_scaled ~ log_popd_1950 | log_popd_1950, data = overlaps_hpd, type = "BC")
# Which is better fit?
lrtest(model1a1.bc, model1a2.bc)
```

```{r}
# Fit model with fixed phi
model1a1.br <- betareg(Percent_overlap_scaled ~ log_popd_1950, data = overlaps_hpd, type = "BR")
# Fit model with variable phi
model1a2.br <- betareg(Percent_overlap_scaled ~ log_popd_1950 | log_popd_1950, data = overlaps_hpd, type = "BR")

# Which is better fit?
lrtest(model1a1.br, model1a2.br)
```

```{r}
## Model outputs
summary(model1a2)
```